Approximation as Differentiation

### 2025-06-09 main/first-order.tex
Minor tweaks.

Following the CHAD approach, automatic differentiation for first-order programs can also be recovered in this
setup, via the total category $\Fam(\FinVect)$. First we recall the chain rule for the forward derivative
$\pushf{f}_x$ and backward derivative $\pullf{f}_x$, which says that derivatives respect
composition~\cite{spivak65}.

@@

\subsubsection{Chain Rule}

Suppose $f: \RR^m \to \RR^n$ and $g: \RR^n \to \RR^k$. Then
for any $x \in \RR^m$ we have:

\begin{itemize}
\item $\pushf{(g \comp f)}_x = \pushf{g}_{f(x)} \comp \pushf{f}_x: \RR^m \to \RR^k$
\item $\pullf{(g \comp f)}_x = \pullf{f}_{x} \comp \pullf{g}_{f(x)}: \RR^k \to \RR^m$
\end{itemize}

@@

We can see these ``optimised'' forward-mode maps as living in $\Fam(\FinVect)$ if we regard every finite
vector space $\RR^n$ as paired with the constant family of tangent spaces at $\RR^n$ (also written $\RR^n$,
for convenience), and each map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as the pair $(f_1, f_2)$.
Then the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp
\partial f)$ again corresponds exactly to the forward chain rule, since $(\reindex{\partial g}{f} \comp
\partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$

>>>>

As per the CHAD work, automatic differentiation for first-order programs can also be recovered in this setup,
using the total category $\Fam(\FinVect)$ instead of $\Fam(\LatGal)$. First we recall the chain rule for
forward derivatives $\pushf{f}_x: \RR^m \linearto \RR^n$ and backward derivatives $\pullf{f}_x: \RR^n
\linearto \RR^m$, which says that derivatives respect composition~\cite{spivak65}.

@@

\begin{proposition}[Chain Rule]
Suppose $f: \RR^m \to \RR^n$ and $g: \RR^n \to \RR^k$ differentiable. Then for any $x \in \RR^m$ we have:

\begin{itemize}
\item $\pushf{(g \comp f)}_x = \pushf{g}_{f(x)} \comp \pushf{f}_x: \RR^m \to \RR^k$
\item $\pullf{(g \comp f)}_x = \pullf{f}_{x} \comp \pullf{g}_{f(x)}: \RR^k \to \RR^m$
\end{itemize}
\end{proposition}

@@

We can see these ``optimised'' forward-mode maps as living in $\Fam(\FinVect)$ if we regard every finite
vector space $\RR^n$ as paired with the constant family of tangent spaces at $\RR^n$ (also written $\RR^n$,
for convenience), and each map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as the pair $(f_1, f_2)$.
Then the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp
\partial f)$ corresponds exactly to the forward chain rule, since $(\reindex{\partial g}{f} \comp \partial
f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$.
